---
title: "Learning Loop Functions"
author: "Godwin Osuji"
date: "12/17/2021"
output: word_document
---
# Simple lapply function

```{r}
x <- list(a = 1:5, b = rnorm(10))
mean_ab <- lapply(x, mean)
mean_ab
```

# Complicated lapply function

```{r}
y <- list(a = 1:5, b = rnorm(10), c = rnorm(20,1), d = rnorm(100,5))
mean_abcd <- lapply(y, mean)
mean_abcd
x <- 1:4
lapply(x, runif)
lapply(x, runif, min = 0, max = 10)
x <- list(a = matrix(1:4, 2,2), b = matrix(1:6, 3,2))
lapply(x, function(elt) elt[,1])
```


# Sapply Function

while **lapply** returns list, **sapply** tries to simplify the result by returning a vector. If there is not way about it then it would return the list.
**runif** returns random numbers

```{r}
x <- 1:4
runif(4, max = 3) #four random numbers with maximum of 3
lapply(x, runif)
lapply(y, mean)
sapply(x, runif)
sapply(y, mean)
```

# Mean of the rows
```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, mean)
rowMeans(matrix(1:10, 5,2))
```

# Mean of the column
```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 2, mean)
colMeans(matrix(1:10, 5,2))  
```

# Using apply to calculate quantile on the rows of a matrix
```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25, 0.75))
```
# Using apply to calculate the average & sum in an array

```{r}
y <- array(rnorm(2*2*10), c(2,2,10))
apply(y, c(1,2,3), mean)
apply(y, 1, mean)
apply(y, c(1,2), sum)
rowMeans(y, dims = 1)
```
# mapply 
**mapply** can be apply on multiple lists unlike **lapply** and **sapply**
```{r}
list(rep(1,4), rep(2,3), rep(3,2), rep(4,1))#instead of this, you could use mapply
mapply(rep, 1:4, c(4,3,2,1)) #Or
mapply(rep, 1:4, 4:1)
noise <-function(n,mean,sd){
  rnorm(n,mean,sd)
}
mapply(noise, 1:5, 1:5, 2) #or
list(noise(1,1,2),noise(2,2,2),noise(3,3,2),noise(4,4,2))
```
# using tapply function
```{r}
x <- c(rnorm(10), runif(10), rnorm(10,1))
y <- gl(6,5)
tapply(x,y,mean)
```
# Split function on a data frame.
The **split function** is actually not a loop function. I need to pay more attention to how the anonymous function is applied here.  

```{r}
s <- split(airquality, airquality$Month) # it converts the Month variable to a factor automatically here.
lapply(s, function(x) colMeans(x[,c("Ozone", "Solar.R", "Wind", "Temp")], na.rm = TRUE))  # more attention on using anonymous function

sapply(s, function(x) colMeans(x[,c("Ozone", "Solar.R", "Wind", "Temp")], na.rm = TRUE))  #it will return a more simplified result; matrix.
```

```{r}
f1<-gl(2,5)
f2<-gl(5,2)
interaction(f1,f2)
x<-rnorm(10)
split(x, list(f1,f2))
split(x, list(f1,f2), drop = TRUE)
```